The Heavy Photon Search Test Detector

### 2014-03-11 hps-testrun-nim.tex
Small edits to trigger section.

The trigger system is designed to select time coincidences of electromagnetic clusters in the 
top and bottom halves of the ECal. Figure~\ref{fig:hps_trigger_cal} shows a schematic overview 
of each stage of the system. 
 \begin{figure}[b]
\begin{center}
{\small
 \includegraphics[width=8cm]{figures/hps_trigger_cal}
\caption{Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes 
signals for each detector channel and sends them for cluster finding in the CTP. The CTP clusters are 
sent to the SSP where the final trigger decision is taken based on pairs of clusters in both halves of the 
ECal. The decision is sent to the Trigger Supervisor (TS) that generates the necessary signals to readout 
the sub-detectors.}
 \label{fig:hps_trigger_cal}
}
\end{center}
 \end{figure}
Each channel on the FADC board have an 
independent data path to send a 5-bit pulse energy and 3-bit pulse arrival time information to a 
trigger processing board (CTP) in the same crate every 32~ns. Contrary to the readout path 
described in Sec.~\ref{sec:fadc}, this energy is a pedestal subtracted time-over-threshold sum with 
programmable offsets and minimum threshold discriminator for each channel. With input from all 
FADC channels, i.e. one half of the ECal, the CTP performs cluster finding and calculates cluster 
energy, shape and timing information. The 3x3 fixed-window, highly parallel, FPGA-based cluster 
algorithm simultaneously searches for up to 125 clusters with energy sum larger than 
the programmable energy threshold (in the Test run this was $\approx270$~MeV). Crystals in the 
fixed-window are included in the sum if the leading edges of the pulses occurred within a 8~ns time 
window (4~ns minimum). To take into account with clock skew and jitter throughout the system 
clustering algorithm can allow crystal hits from $N\in[0,7]$ previous or future 4~ns clock cycles to 
contribute to the cluster sum.  A separate sub-system board (SSP) receives the clusters (only the 
highest energy 3x3 window to deal with overlapping and very large clusters) from 
the top and bottom half CTP every 32~ns and searches for pairs of clusters in a 8~ns wide 
trigger window. The trigger supervisor (TS) generates all the necessary signals and controls the 
entire DAQ system readout through the trigger interface units installed in every crate that participate 
in the readout process.

The trigger system is free-running and driven by the 250~MHz global clock and has essentially zero 
dead time at the occupancies expected for HPS. The Trigger Supervisor can apply dead time if 
necessary, for example on a `busy' or `full' condition from the front-end electronics. The system is 
designed to handle trigger rates above 50~kHz and has a latency set to $\approx 3~\mu$s to match 
that required by the SVT APV25 ASIC. During the Test run, for the most part the trigger system 
required only a single cluster in either the top or bottom Ecal halves and tested to above 100~kHz 
trigger rates with low thresholds.

@@

\subsection{Multiple Coulomb Scattering Distributions}
One of the main goals of the Test run was to evaluate the description of the tails of the 
multiple Coulomb scattering to gain further confidence in the expected detector occupancy in the full HPS 
experiment. Those occupancies have been evaluated using \egs{}~\cite{egs5} simulation program 
whose \moliere{} formulation of multiple scattering show good agreement with previous studies 
for a wide range of materials and projectiles~\cite{Attwood:2005zz,Shen:1978ha,Gottschalk1993467} . 
Despite all data taken with a photon beam in the Test run, the angular distribution of the pair 
produced electron and positron emerging from the converter with different converter thickness can 
be used to study the varying multiple Coulomb scattering contribution (pair production angle is constant). 
Restricting clusters to a well calibrated region of the ECal we observe, within our systematic 
uncertainties , good agreement with the rates predicted by \egs{}, see Tab.~\ref{tab:ang_distr_dataMC}.
%\begin{center}
\begin{table}
{\small
\begin{tabular}{|l|c|c|c|}
\hline
\bf Converter (\%~$X_0$) & \bf 1.60 & \bf 0.45 &	\bf 0.18 \\
\hline
EGS5 &	1162 $\pm$ 112 &	255 $\pm$ 28 &	94 $\pm$ 17	\\
\hline
GEANT4 & 2633 $\pm$ 250 & 	371 $\pm$ 38 &	114 $\pm$ 18 \\
\hline
Observed 	& 1064 $\pm$ 2 & 196 $\pm$ 1 &	92 $\pm$ 1 \\						
\hline
\end{tabular}
\caption{ Observed and predicted number of events for 1~s of beam at 90~nA for three different converter 
thicknesses. The uncertainty on the prediction includes systematic uncertainties from ECal alignment, background normalization, beam current normalization and limited statistics in the simulation.}
}
\label{tab:ang_distr_dataMC}
\end{table}
%\end{center}

>>>>

The trigger system is designed to select time coincidences of electromagnetic clusters in the 
top and bottom halves of the ECal. Figure~\ref{fig:hps_trigger_cal} shows a schematic overview 
of each stage of the system. 
 \begin{figure}[b]
\begin{center}
{\small
 \includegraphics[width=8cm]{figures/hps_trigger_cal}
\caption{Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes 
signals for each detector channel and sends them for cluster finding in the CTP. The CTP clusters are 
sent to the SSP where the final trigger decision is taken based on pairs of clusters in both halves of the 
ECal. The decision is sent to the Trigger Supervisor (TS) that generates the necessary signals to readout 
the sub-detectors.}
 \label{fig:hps_trigger_cal}
}
\end{center}
 \end{figure}
Each channel on the FADC board have an 
independent data path to send 5-bit pulse energy and 3-bit pulse arrival time information every 32~ns 
to a 
trigger processing board (CTP) in the same crate every. Contrary to the readout path 
described in Sec.~\ref{sec:fadc}, this energy is a pedestal subtracted time-over-threshold sum with 
programmable offsets and minimum threshold discriminator for each channel. With input from all 
FADC channels, i.e. one half of the ECal, the CTP performs cluster finding and calculates cluster 
energy, shape and timing information. The 3x3 fixed-window, highly parallel, FPGA-based cluster 
algorithm simultaneously searches for up to 125 clusters with energy sum larger than 
the programmable energy threshold ($\approx270$~MeV). Crystals in the 
fixed-window are included in the sum if the leading edge of the pulse occurred within a 8~ns time 
window (4~ns minimum). To take into account clock skew and jitter throughout the system, the
clustering algorithm can allow for crystal hits from up to seven previous or future 4~ns clock cycles to 
contribute to the cluster sum.  A separate sub-system board (SSP) receives the clusters (only the 
highest energy 3x3 window to deal with overlapping and very large clusters) from 
the top and bottom half CTP every 32~ns and searches for pairs of clusters in a 8~ns wide 
trigger window. The trigger supervisor (TS) generates all the necessary signals and controls the 
entire DAQ system readout through the trigger interface units installed in every crate that participate 
in the readout process.

The trigger system is free-running and driven by the 250~MHz global clock and has essentially zero 
dead time at the occupancies expected for HPS. The trigger supervisor can apply dead time if 
necessary, for example on a `busy' or `full' condition from the front-end electronics. The system is 
designed to handle trigger rates above 50~kHz and has a latency set to $\approx 3~\mu$s to match 
that required by the SVT APV25 ASIC. During the Test run, for the most part the trigger system 
required only a single cluster in either the top or bottom Ecal halves and was tested to above 100~kHz 
trigger rates by lowering thresholds.

@@

\subsection{Multiple Coulomb Scattering Distributions}
\label{sec:mcs}
One of the main goals of the Test run was to evaluate the description of the tails of the 
multiple Coulomb scattering to gain further confidence in the expected detector occupancy in the full HPS 
experiment. Those occupancies have been evaluated using \egs{}~\cite{egs5} simulation program 
whose \moliere{} formulation of multiple scattering show good agreement with previous studies 
for a wide range of materials and projectiles~\cite{Attwood:2005zz,Shen:1978ha,Gottschalk1993467}. 
Despite all data taken with a photon beam in the Test run, the angular distribution of the pair 
produced electron and positron emerging from the converter with different converter thickness can 
be used to study the varying multiple Coulomb scattering contribution (pair production angle is constant). 
Restricting clusters to a well calibrated region of the ECal we observe, within our systematic 
uncertainties , good agreement with the rates predicted by \egs{}, see Tab.~\ref{tab:mcs}.
%\begin{center}
\begin{table}
{\small
\begin{tabular}{|l|c|c|c|}
\hline
\bf Converter (\%~$X_0$) & \bf 1.60 & \bf 0.45 &	\bf 0.18 \\
\hline
EGS5 &	1162 $\pm$ 112 &	255 $\pm$ 28 &	94 $\pm$ 17	\\
\hline
GEANT4 & 2633 $\pm$ 250 & 	371 $\pm$ 38 &	114 $\pm$ 18 \\
\hline
Observed 	& 1064 $\pm$ 2 & 196 $\pm$ 1 &	92 $\pm$ 1 \\						
\hline
\end{tabular}
\caption{ Observed and predicted number of events for 1~s of beam at 90~nA for three different converter 
thicknesses. The uncertainty on the prediction includes systematic uncertainties from ECal alignment, background normalization, beam current normalization and limited statistics in the simulation.}
}
\label{tab:mcs}
\end{table}
%\end{center}
